Deep Reinforcement Learning for Conservation Decisions

### 2021-05-14 manuscript/manuscript.Rmd
Cutting and moving

Among model-free algorithms, there is divergence around what functions the agent is attempting to learn in order to achieve the RL objective.
Generally, model-free agents are either learning a policy function, a value function or both.
For context, *value functions* are proxies for how high of a cumulative reward an agent can expect to receive from a given state or state-action pair.
If an agent knows the value function, then the agent can find the optimal action at any state by selecting the action that will maximize the value function in expectation.
The class of algorithms that exclusively try to learn a value function are called *value learning* methods.
*Policy gradient* algorithms exclusively learn the policy.
And, lastly, *actor-critic* algorithms attempt to learn both a policy function and a value function, whereby the value function, which is learned separately from the policy, is used to inform the agent on the goodness of a selected action.
There is a handful of advantages and disadvantages for each of these approaches that we will not go in depth on -- for more, see [@CITE]; yet, the main point of divergence comes from the bias-variance trade-off [@CITE].
Generally, policy gradient algorithms tend to struggle with high variance, value-based learning tends to be biased while actor-critic methods balance these two concerns [@CITE; @CITE; @CITE].

Neural networks become useful in RL when the environment has a large state-action space, which happens frequently with realistic decision-making problems. 
Classic RL algorithms, like policy iteration, value iteration and TD-learning [@suttonbarto], fail when the state-action space is large, e.g. when the possible states and actions can not be represented in a tractable table [@CITE].
Yet, over the last 10 years, researchers have shown that neural network-based RL algorithms can perform well on previously unsolvable environments.
Neural networks have been widely useful in reinforcement learning because neural nets have the property of being general function approximators [@CITE].
Whenever there is a need for an agent to approximate some function, say a policy function, value function or a transition function, neural networks can be used in this capacity.
While there are other function approximators that can be used in RL, e.g. Gaussian processes, neural networks have excelled in this role because of their ability to learn complex, non-linear, high dimensional functions and their ability to adapt given new information [@CITE].
We will not discuss in detail how neural networks work -- see [@CITE] for further background --, but the general process for how neural networks are used in RL is that at each learning step, the agent, say a policy gradient-based agent, will adjust the parameters in its policy neural network so that according to past experiences, the agent will be more likely to engage in highly rewarding behavior in the future. 
This same process occurs in value learning and actor-critic-based deep RL algorithms, except these agents are adjusting either a value network or both a value and policy network respectively.

@@

## Designing and training an agent
<!-- Maybe rename this as Training and Tuning an Agent -> how to practically get an agent up and running -->

Training a deep, model-free agent on an environment involves allowing the agent to interact with the environment for thousands or even millions of time steps. 
Over the course of training, the agent will engage in a trial and error process whereby the agent continually updates weights on a collection of neural networks in response to the environment's reward signal, much like how reinforcement learning is believed to work in living creatures [@cite].
Fundamental to this trial and error process is a balancing of explorative and exploitative behavior.
The agent will occasionally need to propose exploratory actions that may result in higher rewards than previously seen. 
Yet, the agent will also need to occasionally exploit the best perceived sequence of actions which may require marginal improvement to reach optimality.
As the exploration-exploitation is currently unsolved, there are a range of approaches that RL algorithms employ.

@@

A neural network can be thought of as a generic function approximator, and these weights as parameters of that function that are estimated from 'data' generated by training.
A typical neural network in RL may have thousands or hundreds of thousands of parameters, giving it the flexibility to learn very complex behaviors.
The overall architecture of the neural network, as well as other parameters such as those governing how an agent learns, how an agent proposes actions, and how the agent discounts future rewards, and so forth are controlled by *hyperparameters*.
To initialize a new, untrained agent, we must specify a handful or so of these hyperparameters, which are held fixed over the course of training.
The particular selection of hyperparameters required depends on which *RL algorithm* our agent will be based.
The RL algorithm specifies precisely how the agent will learn from its environment:
Is the algorithm model-based or model-free? 
Stochastic or deterministic? 
"Off-policy" or "on-policy"? 
Does it use a random noise model to propose actions?
Does it permit a continuous action space or require discrete actions?
The design of RL algorithms has been and remains an area of intense research and rapid development [Table 2].
Open source implementations by professional software development teams often follow swiftly after new algorithms are published [@Ssb3; @tf-agents].
In the following section, we summarize some of the key concepts that make up an RL algorithm, and highlight relevant distinctions between important algorithms today.

<!-- Something about given RL algorithm being able to solve lots of different environments with no pre-specified knowledge. cite examples @AlphaZero and @DQN; but still depends on hyperparameters -->
The process of selecting hyperparameters for a given RL algorithm to maximize it's performance on a given environment is referred to as hyperparameter *tuning*.
The process of tuning can be automated through standard multi-dimensional optimization approaches: proposing new combinations of hyperparameters, training an agent, and examining the resulting performance.
Because training even on a fixed set of hyperparameters can be computationally intensive, tuning must be targeted appropriately.
Considerations in how the environment simulation is implemented, such as the normalization of state and/or action spaces, can significantly improve performance of many RL algorithms on a diverse range of problems without requiring extensive tuning.

@@

Many reinforcement learning algorithms center around estimating value functions, which attempt to quantify the goodness of being in a state or taking an action from a state.
Given that the objective of RL is to maximize the cumulative expected reward, "goodness" refers to the likelihood that a state or state-action pair will result in a high cumulative reward trajectory.
There are 3 value functions that are commonly used across RL algorithms: the state-value function, $V^\pi(s_t)$, the state-action value function, $Q^\pi(s_t, a_t)$, and the advantage function, $A^\pi(s_t, a_t)$.
The state-value function is defined as the expected return from a given state,

$$
  V^\pi(s_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

Meanwhile, the state-action-value or quality function, finds the expected return given a state and action,

$$
  Q^\pi(s_t, a_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t, a_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

And lastly, the advantage function characterizes a relative state-action value in that it quantifies how valuable a state-action pair is over a baseline.
The advantage function can then be succinctly defined as

$$
  A^\pi(s_t, a_t) = Q^\pi(s_t, a_t) - V^\pi(s_t)
$$ 
Learning $Q^\pi$, commonly called the Q-value function, is more powerful than learning the state-value function as Q-values easily prescribe what action to take when in a certain state, $a_t^* = \underset{a_t}{\text{argmax}} \, Q^\pi (s_t, a_t)$.
Conversely, if we know the state-value function, we would have to perform a maximization over an expected value to find the optimal action, $a_t^* = \underset{a_t}{\text{argmax}} \, \mathbb{E}_{s_{t+1} \sim \,T(s_{t+1} \mid s_t, a_t)}[V^\pi(s_{t+1})]$, which is less tractable.
Yet, learning the advantage function is often preferred in practice as advantage estimators favorably have lower variance than state-value and Q-value estimators [cite schulman gae].

@@

There is an ever increasing multitude of reinforcement learning algorithms, and while there are far too many algorithms to discuss in detail in this paper, we will herein classify and touch on some current methods.
RL algorithms can generally be split into 2 groups, model-based and model-free RL.
In model-based RL, the agent is attempting to learn the state transition operator, $T$. <!-- Sometimes the agent is just assuming this is given, right? -->
Once the transition model is known, the agent can then use the transition model to plan out what is the optimal action, often with Model Predictive Control [cite tassa 2012 MPC], or the agent can use the transition model to generate more samples to aid learning [cite Sutton dyna, alphago?]. <!-- For most recent methods focus on an approximate solution, since optimal is computationally impossible.  In contrast most readers will assume this step is trivial, if the transition matrix is known there's nothing more to do but "look" at the predicted transitions and make a choice. -->
In model-free RL, the agent does not attempt to learn a dynamics model and instead tries to either learn a parameterized policy, learn a value function, or try to learn both a parameterized policy and a value function.
The general trade-off between model-free and model-based approaches is that in practice, model-based algorithms are often more sample efficient than model-free algorithms but often fail to surpass model-free algorithms in overall agent performance [cite?]

>>>>

@@

## Training and tuning an agent
<!-- Maybe rename this as Training and Tuning an Agent -> how to practically get an agent up and running -->
<!-- Ecologist job to build environment; CS job to build agent 
      Ok so talk about this in very general terms
      selecting an agent, why you might want to use different ones
-->
Training a deep, model-free agent on an environment involves allowing the agent to interact with the environment for thousands or even millions of time steps. 
Over the course of training, the agent will engage in a trial and error process whereby the agent continually updates weights on a collection of neural networks in response to the environment's reward signal, much like how reinforcement learning is believed to work in living creatures [@cite].
Fundamental to this trial and error process is a balancing of explorative and exploitative behavior.
The agent will occasionally need to propose exploratory actions that may result in higher rewards than previously seen. 
Yet, the agent will also need to occasionally exploit the best perceived sequence of actions which may require marginal improvement to reach optimality.
As the exploration-exploitation is currently unsolved, there are a range of approaches that RL algorithms employ.

@@

In practice, the general process for attaining a good performing agent is to tune model hyperparameters then train an agent using the best hyperparameters. 
Hyperparameters refer to the parameters in a RL algorithm that are not directly learned during the training process. <!-- better to define before using in opening sentence.  -->
For example, a weight in the neural network is a model parameter, but the number of hidden layers in the neural network is a hyperparameter.
Since the optimal hyperparameters cannot be found analytically, tuning should be done to find a good-performing set of hyperparameters.
The process of tuning can be automated through standard multi-dimensional optimization approaches: proposing new combinations of hyperparameters, training an agent, and examining the resulting performance.
Since one round of training can be time intensive, hyperparameter tuning, which consists of usually numerous rounds of training, can be particularly time intensive.

@@

@@

There is an ever increasing multitude of reinforcement learning algorithms, and while there are far too many algorithms to discuss in detail in this paper, we will herein classify and touch on some current methods.
